---
title: "Dimension reduction with Flux Balance Analysis"
output: html_notebook
---

# Summary

# 1. Read in libraries
```{r}
library(tidyverse)
library(umap)
library(Seurat)
library(writexl)
```

# 2. Read in data
We'll read in the metabolic
```{r}
rm(list=ls())
flux_data = read_csv('D:/Analysis/Immune/neutrophil_flux.csv')
flux_data[, 2:ncol(flux_data)] = abs(flux_data[, 2:ncol(flux_data)])

# Create identity
labels = flux_data$ID
labels = sub("\\.\\d+$", "", labels)

# Cells need to be cols and rows need to be reactions
flux_data = t(flux_data)

# Set column names
#names(flux_data) = lapply(flux_data[1, ], as.character)
colnames(flux_data) = as.character(flux_data["ID", ])
flux_data = flux_data[-1, ] 
```

Let's create a seurat object
```{r}
neutrophil = CreateSeuratObject(counts=flux_data, 
                                project="neutrophil_flux", 
                                min.cells=1, 
                                min.features=20)
```



# 3. Data Preprocessing
Now that we hacked the Seurat object, let's run the usual preprocessing pipeline. Note that SCTransform doesn't work with this data.

```{r}
neutrophil = FindVariableFeatures(object=neutrophil, 
                              mean.function=ExpMean,
                              dispersion.function=LogVMR,
                              selection.method='vst',
                              x.low.cutoff=-Inf,
                              x.high.cutoff=Inf,
                              y.cutoff=Inf,
                              nfeatures = 2000)

# Log normalize data
neutrophil = NormalizeData(neutrophil, 
                     normalization.method='LogNormalize', 
                     scale.factor=10000)

all.rxns = rownames(neutrophil)
neutrophil = ScaleData(neutrophil, 
                      features=all.rxns)

# Run PCA
neutrophil = RunPCA(object=neutrophil,
              npcs=50,
              verbose=FALSE)

# Find shortest nearest neighbors and clusters
neutrophil = FindNeighbors(neutrophil, 
                     dims=1:50,
                     reduction='pca')

neutrophil = FindClusters(neutrophil, 
                    resolution=0.5)

# Perform UMAP
neutrophil = RunUMAP(neutrophil, 
               dims=1:50)
```

Create the DimPlot.
```{r,fig.height=12}
# Plot based on predicted clusters
DimPlot(neutrophil, 
        reduction="umap",
        dims=c(1, 2),
        label=TRUE,
        repel=TRUE,
        label.size=4,
        na.value = "grey50") + 
        labs(x="UMAP 1", y="UMAP 2", title="Neutrophil metabolic flux UMAP") +
        theme_minimal() +
        theme(plot.title = element_text(size=25),
              legend.text=element_text(size=20),
              axis.title.x=element_text(size=20),
              axis.title.y=element_text(size=20),
              axis.text.x=element_text(size=20),
              axis.text.y=element_text(size=20)) +
        geom_hline(yintercept=0) +
        geom_vline(xintercept=0)
```

Get all metabolic reactions by predicted UMAP clusters
```{r}
cluster_markers = FindAllMarkers(neutrophil,
                         only.pos=TRUE,
                         min.pct=0.25,
                         logfc.threshold=0.25,
                         test.use="MAST",
                         )
cluster_markers = cluster_markers[!duplicated(cluster_markers[, "gene"]), ]

#writexl::write_xlsx(cluster_markers, "D:/Analysis/Immune/neutrophil_flux_clusters.xlsx")
```

Change the identity by day
```{r}
# Create color scheme by changing Identifier
neutrophil.day = SetIdent(obj=neutrophil, 
                          value=labels)

#tmp = AverageExpression(neutrophil.day, features=day_markers$gene)
levels(neutrophil.day) = c("Neutrophils  0", "Neutrophils  7", "Neutrophils  14", "Neutrophils  21")
DoHeatmap(object=neutrophil.day,
          features=day_markers$gene,
          group.colors=c('#E1E5F2', '#BFDBF7', '#1F7A8C', '#022B3A')) +
        scale_y_discrete(labels=day_markers$Name) +
        scale_fill_gradientn(colors=c("blue", "gray", "red"))
        
```

# 4. Run UMAP with the flux data
Rows correspond to different cells in each time point. Columns correspond to different metabolic reactions.
```{r,fig.height=4}
day_markers = FindAllMarkers(neutrophil.day,
                         only.pos=TRUE,
                         min.pct=0.25,
                         logfc.threshold=0.25,
                         test.use="MAST",
                         )

day_markers$Name = c("Adenosine Phosphorylase", 
                         "Adenine Phosphoribosyltransferase",
                         "Adenosine Kinase",
                         "Glutamate Dehydrogenase",
                         "Isocitrate:NAD+ Oxidoreductase",
                         "Isocitrate:NADP+ Oxidoreductase",
                         "Hypoxanthine Phosphoribosyltransferase",
                         "Inosine Phosphorylase",
                         "Inosine Kinase"
                         )

neutrophil_d0 = day_markers[day_markers$cluster=="Neutrophils  0",]$gene
FeaturePlot(neutrophil.day, 
            features=neutrophil_d0[1],
            label=FALSE,
            repel=TRUE,
            label.size=4) + 
        labs(title="Day 0", x="UMAP 1", y="UMAP 2")

neutrophil_d21 = day_markers[day_markers$cluster=="Neutrophils  21",]$gene
FeaturePlot(neutrophil.day, 
            features=neutrophil_d21[1],
            label=FALSE,
            repel=TRUE,
            label.size=4) + 
        labs(title="Day 21", x="UMAP 1", y="UMAP 2")
```
