---
title: "Untitled"
author: "Ryan Rebernick"
date: "3/2/2021"
output: html_document
---

# 1. Load packages and data
```{r, warning=F, message=F}

library('plyr')
library('dplyr')
library('magrittr')
library('data.table')

library(dorothea)
library(Seurat)
library(tibble)
library(pheatmap)
library(tidyr)
library(viper)
library(nichenetr)
library('tibble')
library(ComplexHeatmap)


load('/nfs/turbo/umms-csriram/scampit/Ryan/Projects/Dorothea/data/lung.neut.Robj')

```


# 2. Show existing clustering 
```{r, warning=F, message=F}

# save subtype of neutrophil
neut$neutType <- Idents(neut)

# show clustering
DimPlot(neut, reduction = "tsne")


```

# 3. Get TF activity per cluster using Dorothea
https://saezlab.github.io/dorothea/articles/single_cell_vignette.html
```{r}

# Regulons for mouse
dorothea_regulon_mouse <- get(data("dorothea_mm", package = "dorothea"))

# We obtain the regulons based on interactions with confidence level A, B and C
regulon <- dorothea_regulon_mouse %>%
    dplyr::filter(confidence %in% c("A","B","C"))

# set identities
Idents(neut) <- neut$neutType

# We compute Viper Scores 
neut <- run_viper(neut, regulon,
                  options = list(method = "scale", minsize = 4, 
                                 eset.filter = FALSE, cores = 1, 
                                 verbose = FALSE))

# switch assay to dorothea, scale, run PCA, and find neighbors/clusters
DefaultAssay(object = neut) <- "dorothea"
neut <- ScaleData(neut)
neut <- RunPCA(neut, features = rownames(neut), verbose = FALSE)
neut <- FindNeighbors(neut, dims = 1:10, verbose = FALSE)
neut <- FindClusters(neut, resolution = 0.5, verbose = FALSE)

# Run umap
neut <- RunUMAP(neut, dims = 1:10, umap.method = "uwot", metric = "cosine")

# grab cluster markers
neut.markers <- FindAllMarkers(neut, only.pos = TRUE, min.pct = 0.25, 
                               logfc.threshold = 0.25, verbose = FALSE)


```
# 4. Plot heatmap of TF activity by cluster
## 4A. Prepare dataframe for plotting
The original documentation doesn't filter for factors expressed which makes zero sense to me. I used a mean expression of 0.5 as a cutoff for 'expressed' factors. This significantly narrowd the number of factors down. From 50 predicted ones, to 13 actually expressed.
```{r}

# set identity
Idents(neut) <- neut$neutType

## We transform Viper scores, scaled by seurat, into a data frame to better 
## handling the results
viper_scores_df <- GetAssayData(neut, slot = "scale.data", 
                                    assay = "dorothea") %>%
  data.frame(check.names = F) %>%
  t()

## We create a data frame containing the cells and their clusters
CellsClusters <- data.frame(cell = names(Idents(neut)), 
                            cell_type = as.character(Idents(neut)),
                            check.names = F)

## We create a data frame with the Viper score per cell and its clusters
viper_scores_clusters <- viper_scores_df  %>%
  data.frame() %>% 
  rownames_to_column("cell") %>%
  gather(tf, activity, -cell) %>%
  inner_join(CellsClusters)

## We summarize the Viper scores by cellpopulation
summarized_viper_scores <- viper_scores_clusters %>% 
  dplyr::group_by(tf, cell_type) %>%
  dplyr::summarise(avg = mean(activity),
            std = sd(activity))


## We select the 50 most variable TFs. (200/4pops)
highly_variable_tfs <- summarized_viper_scores %>%
  dplyr::group_by(tf) %>%
  dplyr::mutate(var = var(avg))  %>%
  ungroup() %>%
  top_n(200, var) %>%
  distinct(tf)

## We prepare the data for the plot
summarized_viper_scores_df <- summarized_viper_scores %>%
  semi_join(highly_variable_tfs, by = "tf") %>%
  dplyr::select(-std) %>%   
  spread(tf, avg) %>%
  data.frame(row.names = 1, check.names = FALSE) 

# get highly variable tf mean expression (some factors missing) and compute if TFs expressed (average > 0.5)
Idents(neut) <- neut$orig.ident
cluster.averages <- as.data.frame(AverageExpression(neut)[["RNA"]])
cluster.averages %<>% 
  mutate(gene = rownames(cluster.averages)) %>%
  dplyr::filter(gene %in% highly_variable_tfs$tf) %>%
  mutate(all = as.numeric(all)) %>%
  select(-gene) %>%
  mutate(all = ifelse(all >0.5, 1, 0))

# Filter Dorothea TFs for those expressed
t_summarized_viper_scores_df <- as.data.frame(t(summarized_viper_scores_df))
expressed <- rownames(cluster.averages)[cluster.averages$all>0]
t_summarized_viper_scores_df %<>% dplyr::filter(rownames(.) %in% expressed)


```
## 4B.Plot Dorothea TF activities filtered by TF expression
Used a different heatmap plotting function because the documentation is 1000x better. Long live complexHeatmap
```{r, fig.height=10}

# color palette
palette_length = 100
my_color = colorRampPalette(c("Darkblue", "white","red"))(palette_length)

# create heatmap
h <- Heatmap(t_summarized_viper_scores_df,
        show_column_dend = F,
        show_row_dend = F,
        col = my_color,
        width = unit(1, "in"), 
        height = unit(3, "in"),
        row_names_gp = gpar(fontsize = 8)
        ) 
h

# save to pdf
pdf('/nfs/turbo/umms-csriram/scampit/Ryan/Projects/single-cell-neutrophil-flux-analysis/out/dorothea_neutrophils_expressed_TFs.pdf', height = 5.5, width = 2.5)
h
dev.off()

```

# 5. Write regulons to xlsx document
```{r}
library(xlsx)

Idents(neut) <- neut$neutType

# get expression for all factors
exp <- as.data.frame(AverageExpression(neut)[["RNA"]])
# place holder for append
past1 = F 

for(x in 1:length(expressed)){
print(expressed[x])  
  
# current regulon (TF)
curReg <- regulon %>%
  dplyr::filter(tf == expressed[x])
# expresion of all of the TF targets (regulon)
curExp <- exp %>%
  dplyr::filter(rownames(.) %in% curReg$target)

# save to file
write.xlsx(curExp, file="/nfs/turbo/umms-csriram/scampit/Ryan/Projects/single-cell-neutrophil-flux-analysis/out/S1_regulonExpression.xlsx", sheetName=expressed[x], row.names=T, append = past1)

past1=T

}

```


# 6. Nichenet comparing pro-tumor neutrophils vs anti-tumor/mature neutrophils/MDSCs
#### 6A. combine neutrophil subtype data witih whole lung seurat object
combined with whole lung so can use expression data of other cell types. Ended up specificying that the ligands could come from anywhere. Given that we're studying cancer I think this makes sense as IL1b (the top hit) is known to be secreted by cancer cells and that this influences the microenv. 
```{r}

# load the additional cells to use for signalling if desired
load('/nfs/turbo/umms-csriram/scampit/Ryan/Projects/Dorothea/data/lung.Robj')

# Create metadata with either granulocyte subcluster or cellType info
##all lung cells
subcluster <- as.data.frame(lung$CellType)
colnames(subcluster) <- 'value'
##neutrophil cluster info
toMerge <- as.data.frame(neut$neutType)
colnames(toMerge) <- 'value2'
##merge values
subcluster <- merge(subcluster, toMerge, all.x = T, by = 0)
##reformat and rename PRN
subcluster$value <- as.character(subcluster$value)
subcluster$value2 <-as.character(subcluster$value2)
subcluster$value2 <- ifelse(is.na(subcluster$value2), subcluster$value, subcluster$value2)
rownms <- subcluster$Row.names
subcluster %<>% dplyr::select(value2)
rownames(subcluster) <- rownms

# assign to seurat object
lung$subcluster.celltype <- subcluster
Idents(lung) <- lung$subcluster.celltype

# add additional metadata fields
lung$dxHl <- ifelse(grepl('\\.h\\.', lung$sample), 'Hlthy', 'Dx')


```

#### 6B. Load data required to run nichenet
```{r}
# Ligand target
ligand_target_matrix = readRDS(url("https://zenodo.org/record/3260758/files/ligand_target_matrix.rds"))

# Ligand receptor
lr_network = readRDS(url("https://zenodo.org/record/3260758/files/lr_network.rds"))

# weighted integrated network
weighted_networks = readRDS(url("https://zenodo.org/record/3260758/files/weighted_networks.rds"))

```

#### 6C. Nichenet
Specifically looking at the pro-tumor neutrophils vs all other neutrophils including MDSCs
```{r}

# Add metadata field for comparison
lung$comparison <- ifelse(lung$subcluster.celltype %in% c('anti-tumor neutrophils', 'mature neutrophils', 'MDSCs'), 'anti/mature', lung$subcluster.celltype)

# Run nichnet
Idents(lung) <- lung$CellType
nichenet_output = nichenet_seuratobj_aggregate(
  seurat_obj = lung, 
  receiver = "Neutrophils", 
  condition_colname = "comparison", condition_oi = "pro-tumor neutrophils", condition_reference = "anti/mature", 
  sender = "undefined", 
  ligand_target_matrix = ligand_target_matrix, lr_network = lr_network, weighted_networks = weighted_networks, organism = "mouse")

```

# 7. Plot targets heatmap from Nichenet
```{r}

# plot target heatmap
nichenet_output$ligand_activity_target_heatmap
pdf('/nfs/turbo/umms-csriram/scampit/Ryan/Projects/single-cell-neutrophil-flux-analysis/out/nichenet_neutrophilClusters_sophia.pdf', height = 5.5, width = 7)
nichenet_output$ligand_activity_target_heatmap 
dev.off()


```

# 8. Plot Ligand receptor interactions plot from Nichenet
```{r}

# Receptor interactions
nichenet_output$ligand_receptor_heatmap_bonafide
pdf('/nfs/turbo/umms-csriram/scampit/Ryan/Projects/single-cell-neutrophil-flux-analysis/out/nichenet_neutrophilClusters_receptors_sophia.pdf', height = 3, width = 3.0)
nichenet_output$ligand_receptor_heatmap_bonafide 
dev.off()

```

# 9. Plot expression of ligands across cell types by condition from Nichenet
```{r}

# plot expression of ligands across celltypes by condition
DotPlot(lung, features = nichenet_output$top_ligands %>% rev(), split.by = "dxHl", dot.min = .1) + RotatedAxis()

```

# 10. Plot expression of receptors across neutrophil subtypes from nichenet
```{r}

# plot expression of receptors across neutrophil celltypes by condition
Idents(neut) <- neut$neutType
DefaultAssay(object = neut) <- "RNA" #reset cell object from dorothea
rec <- c('Ifngr1', 'Ifngr2', 'Tnfrsf1a', 'Il6ra', 'Il1r2', 'Il1rap')
DotPlot(neut, features = rec %>% rev(), group.by = "neutType", dot.min = .1) + RotatedAxis()


pdf('/nfs/turbo/umms-csriram/scampit/Ryan/Projects/single-cell-neutrophil-flux-analysis/out/nichenet_neutrophilClusters_receptorExp_sophia.pdf', height = 2, width = 1.75)
DotPlot(neut, features = rec %>% rev(), group.by = "neutType", dot.min = .1, cols = "RdBu") + RotatedAxis() + 
  FontSize(x.text = 8, y.text = 0, y.title = 0, x.title = 0) + NoLegend()
dev.off()

pdf('/nfs/turbo/umms-csriram/scampit/Ryan/Projects/single-cell-neutrophil-flux-analysis/out/nichenet_neutrophilClusters_receptorExp_legend_sophia.pdf', height = 4.5, width = 4.5)
DotPlot(neut, features = rec %>% rev(), group.by = "neutType", dot.min = .1, cols = "RdBu") + RotatedAxis() + 
  FontSize(x.text = 10, y.text = 0, y.title = 0, x.title = 0)  
dev.off()


```


# 11. Plot expression of predicted target genes by cell type from nichenet
```{r}

# get expression for targets 
Idents(neut) <- neut$neutType
targets <- AverageExpression(neut)[["RNA"]]
targets <- targets[rownames(targets) %in% nichenet_output$top_targets,]

# scale and normalize
targets = t(scale(t(targets), center =T, scale=T))

# color palette
palette_length = 100
my_color = colorRampPalette(c("darkorchid4", "white","orange"))(palette_length)


# Plot
H <- Heatmap(targets,
        show_column_dend = F,
        show_row_dend = F,
        col = my_color,
        width = unit(1, "in"), 
        height = unit(2, "in"),
        row_names_gp = gpar(fontsize = 8)
        ) 
H

# save to pdf
pdf('/nfs/turbo/umms-csriram/scampit/Ryan/Projects/single-cell-neutrophil-flux-analysis/out/nichenet_targetGene_scaled_heatmap.pdf', height = 4, width = 2.5)
H
dev.off()


```









